---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

fmir
===
[![Travis build status](https://travis-ci.com/mikmart/fmir.svg?branch=master)](https://travis-ci.com/mikmart/fmir) [![Coverage status](https://codecov.io/gh/mikmart/fmir/branch/master/graph/badge.svg)](https://codecov.io/github/mikmart/fmir?branch=master) [![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)


**fmir** provides simple tools for easily accessing up to date, open weather data from Finland. The data are made available by the [Finnish Meteorological Institute](https://en.ilmatieteenlaitos.fi) and are licensed under [CC-BY 4.0](https://creativecommons.org/licenses/by/4.0/).

Key features in **fmir** include:

* Use simple R syntax to create queries for the FMI API
* Flexibly download XML weather data from the API
* Parse the XML response into a regular data frame

In order to access the download services of the API, you'll need to obtain a key by [registering](https://ilmatieteenlaitos.fi/rekisteroityminen-avoimen-datan-kayttajaksi) for open data use with FMI. See [FMI's open data manual](https://en.ilmatieteenlaitos.fi/open-data) for details.

FMI provides varying types of data in [several different formats](https://en.ilmatieteenlaitos.fi/open-data-manual-fmi-wfs-services) in their API, implemented using OGC Web Feature Service (WFS). Currently the only format supported by **fmir** is the "simple" weather format. For a more comprehensive and feature-rich package for accessing the FMI API, check out https://github.com/rOpenGov/fmi.

## Installation
At the moment, the easiest way to install **fmir** is from github with `devtools`:
```{r install, eval = FALSE}
# install.packages("devtools")
devtools::install_github("mikmart/fmir")
```

## Usage

To get started, set your API key for the session with `fmi_set_key()`:

```{r set-key, eval = FALSE}
library(fmir)
library(ggplot2)

# set your api key for your session
fmi_set_key("insert-your-apikey-here")
```

```{r packages, include = FALSE}
# this README uses an API key from the environment variable FMIR_API_KEY
library(fmir)
library(ggplot2)
```

Once your key is set, you can construct queries to the API with `fmi_query()` and then execute them with `fmi_data()`. A simple query with only a `place` parameter will return weather observations with a 10-minute interval for the past 12 hours:

```{r basic-usage}
# generate a query url with fmi_query
q <- fmi_query(place = c("Espoo", "Oulu", "Rovaniemi"))

# download the data corresponding to your query
weather <- fmi_data(q)
weather

# draw a simple line graph of the recent temperature
ggplot(weather, aes(time, t2m)) + geom_line(aes(colour = place))
```

## In the wild

For a real use-case, check out Pasi Haapakorva's [blog post](https://haapakorva.fi/2018/09/26/2018-09-26-keskil%C3%A4mp%C3%B6tilan-muutos-kuukausittain-oulussa-1955-2018/) (in Finnish) looking at the trend in monthly average temperatures in Oulu in 1955--2018. Spoiler alert: it's been getting warmer.
