---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# fmir
The `fmir` package aims to provide functions for easily accessing up to date, open weather data, made available by the [Finnish Meteorological Institute](https://en.ilmatieteenlaitos.fi). Key features include:

* Use simple R syntax to create queries for the FMI API
* Download machine readable XML weather data from the API
* Parse the XML data to a regular data frame

FMI provides data in several different formats in their API, implemented using OGC Web Feature Service (WFS). Currently the only format supported by `fmir` is the "simple" format.

In order to access the download services of the FMI open data API, you'll need to obtain an API key by registering with the Finnish Meteorological Institute. See [FMI's open data manual](https://en.ilmatieteenlaitos.fi/open-data) for details.

**Note:** `fmir` is still very early in development. Basic functionality is available, but several issues remain: documentation is scarce, and none of the implementation details or even function names can be considered stable.

## Installation
At the moment, the easiest way to install `fmir` is from github with `devtools`:
```{r install, eval = FALSE}
# install.packages("devtools")
devtools::install_github("mikmart/fmir")
```

## Usage
Basic workflow with `fmir` follows these simple steps:
```{r basic-usage, eval = FALSE}
library(fmir)
library(ggplot2)

# set your api key for your session
fmi_set_key("insert-your-apikey-here")

# generate a query url with fmi_query
q <- fmi_query(place = "Oulu")

# download the data corresponding to your query
weather <- fmi_data(q)
head(weather)

# draw a simple line graph of the recent temperature
ggplot(weather, aes(time, t2m)) +
  geom_line()
```
