---
title: "Getting started with fmir"
author: "Mikko Marttila"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

**fmir** provides simple tools for easily accessing up to date, open weather data from Finland. The data are made available by the [Finnish Meteorological Institute](https://en.ilmatieteenlaitos.fi) and are licensed under [CC-BY 4.0](https://creativecommons.org/licenses/by/4.0/).

Key features in **fmir** include:

* Use simple R syntax to create queries for the FMI API
* Flexibly download XML weather data from the API
* Parse the XML response into a regular data frame

In order to access the download services of the API, you'll need to obtain a key by [registering](https://ilmatieteenlaitos.fi/rekisteroityminen-avoimen-datan-kayttajaksi) for open data use with FMI. See [FMI's open data manual](https://en.ilmatieteenlaitos.fi/open-data) for details.

FMI provides varying types of data in [several different formats](https://en.ilmatieteenlaitos.fi/open-data-manual-fmi-wfs-services) in their API, implemented using OGC Web Feature Service (WFS). Currently the only format supported by **fmir** is the "simple" weather format. For a more comprehensive and feature-rich package for accessing the FMI API, check out https://github.com/rOpenGov/fmi.

## Installation

The first step is to install **fmir** from github with `devtools`:
```{r install, eval = FALSE}
# install.packages("devtools")
devtools::install_github("mikmart/fmir")
```

## Usage

The first step is of course to load the package:

```{r packages}
library(fmir)
```

### Setting the API key

To get started with downloading data, set your API key for the session with `fmi_set_key()`:

```{r set-key, eval = FALSE}
fmi_set_key("insert-your-apikey-here")
```

Alternatively, you can also save your API key into an environment variable called `FMIR_API_KEY` for persistence across sessions. That is also the method used for creating this vignette.

### A basic query

Having set your key, we can construct queries to the API with `fmi_query()`, and then execute them with `fmi_data()`. A simple query with only a `place` parameter will return weather observations with a 10-minute interval for the past 12 hours:

```{r basic-usage}
# generate a (vector of) query url(s)
q <- fmi_query(place = c("Espoo", "Oulu", "Rovaniemi"))

# download the data corresponding to the query
weather <- fmi_data(q)
weather
```

Having the data at hand, we can then use **ggplot2** to draw a simple line graph of the recent temperatures in Espoo, Oulu and Rovaniemi:

```{r basic-plot, fig.width=6}
library(ggplot2)

# draw a simple line graph of the recent temperature
ggplot(weather, aes(time, t2m)) + geom_line(aes(colour = place))
```
